Evaluation of pilot jobs for Apache Spark\\ applications on HPC clusters

### 2019-05-17 paper.tex
updated figures and added text on spark

@@

        \todo{scala standalone vs our pyspark workaround. kill masters in experiments}
    % maybe remove\subsection{Checkpointing}
        \todo{metric for determining how often to checkpoint based on cluster size}
    % maybe remove \subsection{Job arrays}
        \todo{need to kill idle workers. may not want all workers to be running at once.}
    % maybe remove \subsection{Example application}
        \todo{incrementation with varying task durations}

@@

In the case of pilots, multiple Slurm jobs, each with potentially different queuing times,
are used. Figure~\ref{fig:mwall} displays the makespan times for an average number of workers
given an experiment configuration, where the line corresponds to the estimated makespan using 
our model. Makespan for the model was calculated as:
$$
M = (\lceil 125/W \rceil\times d) + 20,
$$
where:
\begin{itemize}
    \item $M$ is the makespan of the application
    \item $W$ is the average number of Spark workers throughout the execution
    \item $d$ is the task delay associated with a given configuration
    \item $20$ is the average measured incrementation duration for a BigBrain chunk
\end{itemize}
\TG{You should remove the $\lceil . \righceil$ as it is not part of the model and I'm not sure
if it works with pilots. Also, if you're using numbers you should report their units (seconds here, for 20 and the makespan).
It's unclear where 125 comes from. And why isn't it $\frac{125(d+20)}{W}$?}
As can be seen in the figure, although pilots and batch had similar makespans
for a given number of average workers, pilots were consistently slower than batch
for the same number of workers. Both batch and pilots were consistent with the 
model, however \TG{there is a contradiction here: if both batch and pilots were consistent with the model, then
they should have similar makespans.}. In general, both pilots and batch had the same number of average
workers. \TG{You could also plot the difference in the number of workers, for each experiment.}

@@

@@

Although queuing time is largely responsible for makespan variations, it is not
always entirely responsible for the difference. Sometimes there are errors related
to workers not registering properly with the master. These occurrences lead to the 
application being processed entirely with a smaller number of available workers. As the task
delay added considers the maximum amount of parallel workers, functioning with less total workers
will significantly affect makespan. However, as seen in Figure~\ref{fig:mwall}
pilots are slower even with the same amount of average workers. A potential reasoning for this
may be the worker registration delay and the time required to transfer data
over after the application has already started. In batch scheduling, all workers are started in
parallel and the data is transferred automatically at the beginning of the pipeline as the maximum amount
of parallelism is available at driver start time. Pilots, on the other hand, may start workers in parallel,
but not necessarily all will be started at the same time. This means that the entire 
workflow may suffer the impacts of starting workers in sequence \TG{isn't it queuing time again?}. Moreover, there would also
be a delay with respect to transferring data from workers to newly added workers. Therefore,
it is only natural that pilots would have a bit more overhead that batch given the same
queuing time should pilots not all commence at once \TG{ok, but isn't the total amount of transfered data the same in both conditions?}.
 Furthermore, the pilot scripts also
have some starter overhead as the all \TG{?} start a master, a worker (registered to the main
master) and attempt to start a driver. Batch needs only start a single master and attempts to start
a driver only once.

>>>>

    \section{Spark Configuration}
    Spark has many configuration options available for dynamic resource provisioning.
    Its standalone cluster provides two deployment modes: client and cluster. The client
    deploy mode executes the driver within the \texttt{spark-submit} process as
    a client to the cluster. The cluster deploy-mode, however, runs the driver within
    a worker process. Such a configuration is practical when the driver program
    is deployed on a machine not co-located in the cluster to reduce network 
    latency between workers and driver. The cluster deploy mode also has the texttt{supervise}
    feature which provides fault-tolerance to the driver. This is particularly useful for
    pilot-based overlay clusters where the running driver may fail to walltime expiration.
    Cluster deploy mode is not available for python applications using the standalone mode.

    With HPC clusters, client mode is not necessarily possible on the login node as the
    driver client may require more resources than permitted. Moreover, due to cluster
    security policies, it may not be possible, or difficult to execute the driver from
    an external computer, such as a personal laptop. Therefore, for both batch and pilot
    scheduling, it is necessary to launch the driver from within the Slurm jobs.

    Spark also provides master fault tolerance through single-node recovery, where a shared
    filesystem is available. To enable this, a recovery folder must be set. Should the master
    fail, a new master will be able to takeover based on the information available in the recovery
    folder. Although we start a master on each pilot, we have not incorporated this recovery function
    for our experiments. This feature is, however, important and should be incorporated in pilot schedulers
    were any nodes can fail by design.

    %%maybe add checkpointing

@@

    % maybe remove\subsection{Checkpointing}
    %    \todo{metric for determining how often to checkpoint based on cluster size}
    % maybe remove \subsection{Job arrays}
    %    \todo{need to kill idle workers. may not want all workers to be running at once.}
    % maybe remove \subsection{Example application}
    %    \todo{incrementation with varying task durations}

@@

@@

In the case of pilots, multiple Slurm jobs, each with potentially different queuing times,
are used. Figure~\ref{fig:mwall} displays the makespan times for an average number of workers
given an experiment configuration, where the line corresponds to the estimated makespan using 
our model. Makespan for the model was calculated as:
$$
M = \frac{125\times(d + 20)}{W}\times 10,
$$
where:
\begin{itemize}
    \item $125$ is the number of chunks BigBrain was split into
    \item $M$ is the makespan of the application
    \item $W$ is the average number of Spark workers throughout the execution
    \item $d$ is the task delay associated with a given configuration
    \item $20$ is the average measured incrementation duration, in seconds, for a BigBrain chunk
    \item $10$ is the number of iterations
\end{itemize}
As can be seen in the figure, although pilots and batch had similar makespans
for a given number of average workers, pilots were consistently slower than batch
for the same number of workers. Both batch and pilots were consistent with the 
model, however \TG{there is a contradiction here: if both batch and pilots were consistent with the model, then
they should have similar makespans.}. In general, both pilots and batch had the same number of average
workers. \TG{You could also plot the difference in the number of workers, for each experiment.}

@@

Although queuing time is largely responsible for makespan variations, it is not
always entirely responsible for the difference. Sometimes there are errors related
to workers not registering properly with the master. These occurrences lead to the 
application being processed entirely with a smaller number of available workers. As the task
delay added considers the maximum amount of parallel workers, functioning with less total workers
will significantly affect makespan. However, as seen in Figure~\ref{fig:mwall}
pilots are slower even with the same amount of average workers. A potential reasoning for this
may be the worker registration delay and the time required to transfer data
over after the application has already started. In batch scheduling, all workers are started in
parallel and the data is transferred automatically at the beginning of the pipeline as the maximum amount
of parallelism is available at driver start time. Pilots, on the other hand, may start workers in parallel,
but not necessarily all will be started at the same time. This means that the entire 
workflow may suffer the impacts of starting workers in sequence \TG{isn't it queuing time again?}. Moreover, there would also
be a delay with respect to transferring data from workers to newly added workers. Therefore,
it is only natural that pilots would have a bit more overhead than batch given the same
queuing time should pilots not all commence at once \TG{ok, but isn't the total amount of transfered data the same in both conditions?}.
 Furthermore, the pilot scripts also
have some startup overhead as all pilots start a master, a worker (registered to the main
master) and attempt to start a driver. Batch needs only start a single master and attempts to start
a driver only once.
